---
title: "4_ICA_mw_32and36R"
author: "Jackie"
date: "5/12/2021; 6/28/21"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

Computing inter-network functional connectivity

"partial correlation analysis among these 4 ICs using their associated time courses, which resulted in a 4 × 4 symmetric intrinsic network connectivity (INC) matrix for each participant. Each of the 16 unique cells in the matrix reflects the inter-network Fc between a pair of ICs" (Li et al., 2017, _Journal of Psychiatric Research_)

"Then normalize with Fisher r to z transform" (Cao et al., 2020, _Psychiatry Research_)

# loading libraries
```{r, message=FALSE}
library(lmerTest)
library(tidyverse)
library(foreign)
library(haven)
library(readxl)
library(naniar)
library(lubridate)
library(psych)
library(corrr)
library(corrplot)
library(gvlma)
library(mgcv)
library(sjmisc)
library(sjPlot)
library(car)
library(BayesFactor)
```

# Reading in mw and puberty/demo dfs

```{r filepaths}
mw_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/0_MW_Act_Demo_Descriptives/MW_daily_sleep_data_reduced.csv"

# full AHC sample with pubertal and demographic info
pub_demo_act_mw_comb_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/0_MW_Act_Demo_Descriptives/dem_and_tanner_mw_act_comb_merge.csv"
```

```{r, message=FALSE}
mw <- read_csv(mw_fp) %>%
  mutate(ELS_ID = factor(ELS_ID)) %>%
  as_tibble()

pub <- read_csv(pub_demo_act_mw_comb_fp) %>% # remember this dataframe is based on a bunch of scriptss that aligned the tanner and age with the AHC date (all variables - eg, Sex, BMI are based on the AHC timepoint)
  mutate(ELS_ID = factor(ELS_ID)) %>%
  as_tibble()
```

## filtering for EMA
```{r}
pub_filt <-
  pub %>%
  filter(
    !is.na(MW_daily_sleep_session_usable)
  )
```

## merging dfs
And figuring out if any rows were duplicated by merging
```{r}
pub_mw <-
  left_join(
    pub_filt,
    mw,
    by = "ELS_ID"
  ) %>%
  mutate(
    dup = duplicated(ELS_ID)
  )

ELSdup <- pub_mw %>% filter(dup == TRUE) # none

write_csv(pub_mw, "pub_mw_merged.csv")
```


# reading in intra- and inter-network coherence values, demographics, mean fd
```{r, message=FALSE}

mw_36R_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/3_study3/GICAandDualReg_MW36R/coherence_mw_36R.xlsx"
mw_36R <- read_excel(mw_36R_fp, sheet = "comb_coherence_mw_36R") %>%
  mutate(ELS_ID = factor(ELS_ID)) %>%
  mutate(
    timepoint_mw_36R =
      ifelse(
        str_detect(ELS_ID, "ses-T3x"),
        "T3x",
        "T3"
      ),
    timepoint_mw_36R =
      ifelse(
        str_detect(ELS_ID, "ses-T4"),
        "T4",
        timepoint_mw_36R
      )
  ) %>%
  relocate(timepoint_mw_36R, .after = ELS_ID)

inter_net_corr_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/3_study3/net2net/inter_net_corr_merged_mw_36R.csv"
inter_net <- read_csv(inter_net_corr_fp) %>%
  mutate(ELS_ID = factor(ELS_ID))

meanfd_and_vols_t3_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/3_study3/meanFD_ELS-T3_T3x.csv"
meanfd_and_vols_t3 <- read_csv(meanfd_and_vols_t3_fp) %>%
  mutate(ELS_ID = factor(ELS_ID)) %>%
  dplyr::select(ELS_ID, meanFD) %>%
  rename(
    meanFD_t3 = meanFD
  )
meanfd_and_vols_t4_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/3_study3/meanFD_ELS-T4_T4x.csv"
meanfd_and_vols_t4 <- read_csv(meanfd_and_vols_t4_fp) %>%
  mutate(ELS_ID = factor(ELS_ID)) %>%
  dplyr::select(ELS_ID, meanFD) %>%
  rename(
    meanFD_t4 = meanFD
  )

sensitivity_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/SacheCoury/Jackie_dissertation/assessing_numbers/els_subjs_list_rest_ica_ahc_excludingthosewithlargeinterval.csv"
sensitivity <- read_csv(sensitivity_fp)


scan_covariates_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/3_study3/scan_covariates.csv"
scan_covariates <- read_csv(scan_covariates_fp) %>%
  drop_na(mri_survey_timestamp) %>%
  mutate(
    ELS_ID = factor(ELS_ID)
  )

cdi_t3t4_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/3_study3/cdi_t3t4.csv"
cdi_t3t4 <- read_csv(cdi_t3t4_fp) %>%
  mutate(
    ELS_ID = factor(ELS_ID)
  ) %>%
  dplyr::select(ELS_ID, Timepoint_cdi.T3, cdi_total.T3, Timepoint_cdi.T4, cdi_total.T4)

siq_t3t4_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/3_study3/siq_t3t4.csv"
siq_t3t4 <- read_csv(siq_t3t4_fp) %>%
  mutate(
    ELS_ID = factor(ELS_ID)
  ) %>%
  dplyr::select(ELS_ID, Timepoint_siq.T3, siq_total.T3, Timepoint_siq.T4, siq_total.T4)
```

### functions
```{r}
getmode <- function(v, na.rm=TRUE) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

```

## merging
```{r}
meanfd <-
  full_join(
    meanfd_and_vols_t3,
    meanfd_and_vols_t4,
    by = "ELS_ID"
  ) %>%
  mutate(
    ELS_ID = str_remove(ELS_ID, "sub-"),
    ELS_ID = as.numeric(ELS_ID),
    ELS_ID = factor(ELS_ID)
  ) %>%
  filter(!(duplicated(ELS_ID)))

dup <-
  meanfd %>%
  filter(duplicated(ELS_ID)) # 0 dups

sensitivity_clean <-
  sensitivity %>%
  mutate(
    ELS_ID = str_remove(ELS_ID, "sub-"),
    ELS_ID = as.numeric(ELS_ID),
    ELS_ID = factor(ELS_ID)
  )   
```

### variable names
```{r}
vars_36R <- c("aDMN_Z", "lECN_Z", "Noise_brainstem_Z", "pDMN_Z", "rECN_Z")
```

### making ELS_ID compatible data type for merging with other dataframes
```{r}
# for 36R
mw_36R_clean <-
  mw_36R %>%
  mutate(
    timepoint =
      ifelse(
        str_detect(ELS_ID, "T3"),
        "T3",
        "T4"
        ),
    ELS_ID_shortened = 
      substr(ELS_ID, 5, 7)
    ) %>%
  relocate(ELS_ID_shortened, .after = ELS_ID) %>%
  relocate(timepoint, .after = ELS_ID_shortened) %>%
  mutate(
    ELS_ID_shortened = 
      as.numeric(ELS_ID_shortened),
    ELS_ID_shortened =
      factor(ELS_ID_shortened)
    ) %>%
  dplyr::select(-ELS_ID) %>%
  rename(ELS_ID = ELS_ID_shortened) %>%
  dplyr::select(-salience_Z, -thalamicstriatal_Z)

mw_36R_clean_tmpt <-
  mw_36R_clean %>%
  rename(
    timepoint_36R = timepoint,
    aDMN_Z_36R = aDMN_Z,
    lECN_Z_36R = lECN_Z,
    Noise_brainstem_Z_36R = Noise_brainstem_Z,
    pDMN_Z_36R = pDMN_Z,
    rECN_Z_36R = rECN_Z
  ) %>%
  dplyr::select(-timepoint_36R)
```


## merging with scan info
```{r}
mw_36R_clean_scaninfo <-
  left_join(
    mw_36R_clean_tmpt,
    scan_covariates,
    by = "ELS_ID"
  ) %>%
  rename(
    scan_timepoint = timepoint
  ) %>%
  relocate(scan_timepoint, .after = timepoint_mw_36R) %>%
  mutate(
    drop =
      ifelse(
        timepoint_mw_36R == scan_timepoint,
        "keep",
        "drop"
      )
  ) %>%
    relocate(drop, .after = scan_timepoint)

mw_36R_clean_scaninfo_keep <-
  mw_36R_clean_scaninfo %>%
  filter(
    drop == "keep"
  )
```

# excluding those with scan way far back
```{r}
mw_36R_clean_sens <-
  left_join(
    mw_36R_clean_scaninfo_keep,
    sensitivity_clean,
    by = "ELS_ID"
  ) %>%
  filter(
    drop_for_mw == "keep"
  ) %>%
  mutate(
    mw_and_scan_align =
      ifelse(
        usable_MW_DailySleep == restingstate_timept |
          restingstate_timept == "BOTH",
        "yes",
        "no"
      )
  ) %>%
  filter(mw_and_scan_align == "yes") %>%
  filter(usable_MW_DailySleep != "NOT USABLE")

write_csv(mw_36R_clean_sens, "mw_36R_clean_sens.csv")
```

## distributions of networks with mw sample 36 R
```{r}
admn_hist_36R <- 
  mw_36R_clean_sens %>%
  ggplot(
    aes(x = aDMN_Z_36R)
    ) +
  geom_histogram(fill="firebrick3", alpha = 1) +
  theme_classic() +
  labs(
    x = "anterior DMN within-network FC"
  )
admn_hist_36R
ggsave("admn_hist_mw_36R.png", admn_hist_36R)

pdmn_hist_36R <-
  mw_36R_clean_sens %>%
  ggplot(
    aes(x = pDMN_Z_36R)
    ) +
  geom_histogram(fill="mediumblue", alpha = 1) +
  theme_classic() +
  labs(
    x = "posterior DMN within-network FC"
  )
pdmn_hist_36R
ggsave("pdmn_hist_mw_36R.png", pdmn_hist_36R)

recn_hist_36R <-
  mw_36R_clean_sens %>%
  ggplot(
    aes(x = rECN_Z_36R)
    ) +
  geom_histogram(fill="gold", alpha = 1) +
  theme_classic() +
  labs(
    x = "right FPN within-network FC"
  )
recn_hist_36R
ggsave("recn_hist_mw_36R.png", recn_hist_36R)

lecn_hist_36R <-
  mw_36R_clean_sens %>%
  ggplot(
    aes(x = lECN_Z_36R)
    ) +
  geom_histogram(fill="olivedrab3", alpha = 1) +
  theme_classic() +
  labs(
    x = "left FPN within-network FC"
  )
lecn_hist_36R
ggsave("lecn_hist_mw_36R.png", lecn_hist_36R)

noise_brainstem_hist_36R <-
  mw_36R_clean_sens %>%
  ggplot(
    aes(x = Noise_brainstem_Z_36R)
    ) +
  geom_histogram(fill="grey60", alpha = 1) +
  theme_classic() +
  labs(
    x = "Brainstem within-network FC"
  )
noise_brainstem_hist_36R 
ggsave("noise_brainstem_hist_mw_36R.png", noise_brainstem_hist_36R)

library(cowplot)
all_hist_mw_36R <- 
  cowplot::plot_grid(
  admn_hist_36R,
  pdmn_hist_36R,
  lecn_hist_36R,
  recn_hist_36R,
  noise_brainstem_hist_36R,
  labels = c('A',
             'B',
             'C',
             'D',
             'E'),
  label_size = 10,
  ncol = 2
  )
ggsave("all_hist_mw_36R.png", plot = all_hist_mw_36R, height = 7, width = 7)
```

# creating correlation matrix
```{r}
mw_36R_clean_cor <-
  mw_36R_clean_sens %>%
  dplyr::select(-c(ELS_ID, timepoint_mw_36R, scan_timepoint, starts_with("drop"), mri_survey_timestamp, mri_survey_timestamp_recoded, hourofscan, mrisurv_caffeine_day_of_scan, mrisurv_sleepiness_during_scan, mrisurv_sleep_med, starts_with("mrisurv_med_"), usable_MW_DailySleep, usable_Actiwatch, restingstate_timept, ICA_AHC, mw_and_scan_align)) %>%
  drop_na(mrisurv_sleep_before_scan_usual)
cor_mat <- cor(mw_36R_clean_cor, method = "pearson")
corrplot(cor_mat, addCoef.col = TRUE, method = "square", type = "lower", pch.col = "black", tl.col = "black")

```


# computing/describing demographic info
```{r}
dem_plus_mw <-
  left_join(
    mw_36R_clean_sens,
    pub_mw,
    by = "ELS_ID"
  ) 
dem_plus_mw <-
  left_join(
    dem_plus_mw,
    cdi_t3t4,
    by = "ELS_ID"
  )
dem_plus_mw <-
  left_join(
    dem_plus_mw,
    siq_t3t4,
    by = "ELS_ID"
  )
dem_plus_mw <-
  left_join(
    dem_plus_mw,
    meanfd,
    by = "ELS_ID"
  )

dem_plus_mw_clin <-
  dem_plus_mw %>%
  mutate(
    cdi_total_with_mw =
      ifelse(
        str_detect(timepoint_mw_36R, "T3"),
        cdi_total.T3,
        NA
      ),
    cdi_total_with_mw =
      ifelse(
        str_detect(timepoint_mw_36R, "T4"),
        cdi_total.T4,
        cdi_total_with_mw
      ),
    siq_total_with_mw =
      ifelse(
        str_detect(timepoint_mw_36R, "T3"),
        siq_total.T3,
        NA
      ),
    siq_total_with_mw =
      ifelse(
        str_detect(timepoint_mw_36R, "T4"),
        siq_total.T4,
        siq_total_with_mw
      )    
    ) %>%
    drop_na(dailysleep_sat_mean) %>%
  mutate(
    Child_Sex.T1_rec = factor(Child_Sex.T1_rec),
    KSADS_Child_Race_by_P.T1 = factor(KSADS_Child_Race_by_P.T1),
    meanFD =
      ifelse(
        timepoint_mw_36R == "T3",
        meanFD_t3,
        meanFD_t4
      )
  ) # one had no sleep sat score

dem_plus_mw_clindup <-
  dem_plus_mw_clin %>%
  filter(duplicated(ELS_ID)) # no dup

sex_count <-
  dem_plus_mw_clin %>%
  group_by(Sex) %>%
  summarize(n = n()) 
# 30 Male, 43 Female
```


# merging inter-network correlations
```{r}
inter_net <-
  inter_net %>%
  mutate(
    ELS_ID = str_remove(ELS_ID, "sub-")
  ) %>%
  mutate(ELS_ID = as.numeric(ELS_ID),
         ELS_ID = factor(ELS_ID))

dem_plus_mw_clin_netcorr <-
  left_join(
    dem_plus_mw_clin,
    inter_net,
    by = "ELS_ID"
    )

dem_plus_mw_clin_netcorr <-
  dem_plus_mw_clin_netcorr %>%
  mutate_at(
    vars(Gender,mrisurv_caffeine_day_of_scan, mrisurv_sleep_before_scan_usual, mrisurv_sleepiness_during_scan, mrisurv_sleep_med, mrisurv_med_general), factor) %>%
 mutate(
    interval_btwn_mw_and_scan =
    ifelse(
      mri_survey_timestamp_recoded < mw_start_date,
      (mri_survey_timestamp_recoded %--% mw_start_date)/ddays(1),
      (mw_start_date %--% mri_survey_timestamp_recoded)/ddays(1)
      ),
    interval_btwn_scan_tanner =
    ifelse(
      mri_survey_timestamp_recoded < tanner_date_for_mw,
      (mri_survey_timestamp_recoded %--% tanner_date_for_mw)/ddays(1),
      (tanner_date_for_mw %--% mri_survey_timestamp_recoded)/ddays(1)
      ) 
    )

```

# demographics
```{r}
dem_mw_scan <-
  dem_plus_mw_clin_netcorr %>%
  summarize(
    n = n(),
    n_female = sum(Sex == 2),
    n_female_perc = (n_female/n)*100,
    n_gender = sum(Gender == 2),
    n_gender_perc = (n_gender/n)*100,
    n_covid = sum(COVID_AHC_MW == 1),
    n_covid_perc = (n_covid/n)*100,
    n_caffeine_dayofscan = sum(mrisurv_caffeine_day_of_scan == 1, na.rm = TRUE),
    n_caffeine_dayofscan_perc = (n_caffeine_dayofscan/n)*100,
    n_sleepiness_during_scan = sum(mrisurv_sleepiness_during_scan == 1, na.rm = TRUE),
    n_sleepiness_during_scan_perc = (n_sleepiness_during_scan/n)*100,
    n_sleepmed = sum(mrisurv_sleep_med == 1, na.rm = TRUE),
    n_sleepmed_perc = (n_sleepmed/n)*100,
    n_med_general = sum(mrisurv_med_general == 1, na.rm = TRUE),
    n_med_general_perc = (n_med_general/n)*100,
    age_mean = mean(tanner_age_for_mw),
    age_sd = sd(tanner_age_for_mw),
    age_min = min(tanner_age_for_mw),
    age_max = max(tanner_age_for_mw),
    tanner_mean = mean(tanner_average_for_mw),
    tanner_sd = sd(tanner_average_for_mw),
    tanner_min = min(tanner_average_for_mw),
    tanner_max = max(tanner_average_for_mw),
    bmi_mean = mean(BMI_MW, na.rm = TRUE),
    bmi_sd = sd(BMI_MW, na.rm = TRUE),
    bmi_min = min(BMI_MW, na.rm = TRUE),
    bmi_max = max(BMI_MW, na.rm = TRUE),
    householdincome_mean = mean(Household_Income_MW, na.rm = TRUE),
    householdincome_sd = sd(Household_Income_MW, na.rm = TRUE),
    householdincome_min = min(Household_Income_MW, na.rm = TRUE),
    householdincome_max = max(Household_Income_MW, na.rm = TRUE),
    householdincome_mode = getmode(Household_Income_MW),
    sumsevtype_mean = mean(sumsev_type),
    sumsevtype_sd = sd(sumsev_type),
    sumsevtype_min = min(sumsev_type),
    sumsevtype_max = max(sumsev_type),
    intervalbtwn_mw_and_scan_mean = mean(interval_btwn_mw_and_scan),
    intervalbtwn_mw_and_scan_sd = sd(interval_btwn_mw_and_scan),
    intervalbtwn_mw_and_scan_min = min(interval_btwn_mw_and_scan),
    intervalbtwn_mw_and_scan_max = max(interval_btwn_mw_and_scan),
    meanfd_mean = mean(meanFD),
    meanfd_sd = sd(meanFD),
    meanfd_min = min(meanFD),
    meanfd_max = max(meanFD),
    asq_mean = mean(asq_total_mw, na.rm = TRUE),
    asq_sd = sd(asq_total_mw, na.rm = TRUE),
    asq_min = min(asq_total_mw, na.rm = TRUE),
    asq_max = max(asq_total_mw, na.rm = TRUE),      
    race_count_wh = sum(KSADS_Child_Race_by_P.T1 == "1"),
    race_count_bla = sum(KSADS_Child_Race_by_P.T1 == "2"),
    race_count_lat = sum(KSADS_Child_Race_by_P.T1 == "3"),
    race_count_as = sum(KSADS_Child_Race_by_P.T1 == "4"),
    race_count_bir = sum(KSADS_Child_Race_by_P.T1 == "5"),
    race_count_oth = sum(KSADS_Child_Race_by_P.T1 == "6"),
    race_perc_wh = (race_count_wh/n)*100,
    race_perc_bla = (race_count_bla/n)*100,
    race_perc_lat = (race_count_lat/n)*100,
    race_perc_as = (race_count_as/n)*100,
    race_perc_bir = (race_count_bir/n)*100,
    race_perc_oth = (race_count_oth/n)*100,
  )
write_csv(dem_mw_scan, "demo_table_mw_scan.csv")
```

# within-network coherence and sleep satisfaction
```{r}
# sleep sat outliers
sleepout = boxplot(dem_plus_mw_clin_netcorr$dailysleep_sat_mean)$out 
# no outliers
```

## centering
I standardized tanner stage, age at tanner assessment, and bmi within each sex
```{r}
dem_plus_mw_clin_cent <-
  dem_plus_mw_clin_netcorr %>%
  mutate(
    Sex = factor(Sex),
    mw_dailysleepsat_z = scale(dailysleep_sat_mean),
    mw_dailysleephrs_z = scale(dailysleep_hrs_mean),
    COVID_AHC_ACT = factor(COVID_AHC_ACT),
    aDMN_Z_36R_z = scale(aDMN_Z_36R),
    pDMN_Z_36R_z = scale(pDMN_Z_36R),
    admn_lecn_mw_z = scale(admn_lecn_mw),
    pdmn_lecn_mw_z = scale(pdmn_lecn_mw),
    admn_recn_mw_z = scale(admn_recn_mw),
    pdmn_recn_mw_z = scale(pdmn_recn_mw),
    Noise_brainstem_Z_36R_z = scale(Noise_brainstem_Z_36R) ,
    asq_total_mw_z = scale(asq_total_mw)
    ) %>%
  group_by(Sex) %>%
  mutate(
    tanner_avg_at_mw_z = scale(tanner_average_for_mw), # standardize within each sex
    tanner_adrenal_mw_z = scale(tanner_adrenal_mw),
    tanner_gonadal_mw_z = scale(tanner_gonadal_mw),
    bmi_at_mw_z = scale(BMI_MW),
    age_at_tanner_at_mw_z = scale(tanner_age_for_mw)
  ) %>%
  ungroup()
```


# test of potential covariates
```{r}
# corr btwn sleep sat and interval between mw and scan
cor.test(dem_plus_mw_clin_cent$mw_dailysleepsat_z, dem_plus_mw_clin_cent$interval_btwn_mw_and_scan) # no corr, r = -0.1197761, p = 0.3233

# corr btwn sleep sat and mean FD
cor.test(dem_plus_mw_clin_cent$mw_dailysleepsat_z, dem_plus_mw_clin_cent$meanFD) # no corr, r = 0.1809622, p = 0.1338

# corr btwn sleep sat and age
cor.test(dem_plus_mw_clin_cent$mw_dailysleepsat_z, dem_plus_mw_clin_cent$age_at_tanner_at_mw_z)  # no corr, r = -0.05199585 , p = 0.669 (as seen in our puberty analysis)

# corr btwn sleep sat and tanner
cor.test(dem_plus_mw_clin_cent$mw_dailysleepsat_z, dem_plus_mw_clin_cent$tanner_avg_at_mw_z) # yes, corr, r = -0.260141, p = 0.02964 (as seen in our puberty analysis)

# association btwn sleep sat and sex
summary(lm(mw_dailysleepsat_z ~ Sex, data = dem_plus_mw_clin_cent)) # no association, B = -0.2904, p = 0.234

# corr btwn sleep sat and dperession severity
cor.test(dem_plus_mw_clin_cent$mw_dailysleepsat_z, dem_plus_mw_clin_cent$cdi_total_with_mw)  # no corr, r = -0.2172398, p = 0.07516

# corr btwn sleep sat and SI severity
cor.test(dem_plus_mw_clin_cent$mw_dailysleepsat_z, dem_plus_mw_clin_cent$siq_total_with_mw)  # no corr, r = -0.2427267, p = 0.05329)

# corr btwn sleep sat and perceived stress
cor.test(dem_plus_mw_clin_cent$mw_dailysleepsat_z, dem_plus_mw_clin_cent$asq_total_mw)  # yes corr, r = -0.2761626, p = 0.02846
```

```{r}
# distribution of time of scan
timeofscanplot <- 
  dem_plus_mw_clin_cent %>%
  ggplot(
    aes(x=hourofscan)
  ) +
  geom_freqpoly() + 
  theme_classic() +
  labs(
    x = "Time of Scan",
    title = "Distribution of Scan Time of Day"
    ) +
  scale_x_time(breaks = scales::breaks_width("2 hours"))
timeofscanplot
ggsave("timeofscanplot_mw.png", timeofscanplot, width = 7, height = 5)
```


# associations between inter-network connectivity and sleep
```{r}
options(scipen = 999)

m1 <- lm(mw_dailysleepsat_z ~ admn_lecn_mw_z + pdmn_lecn_mw_z + tanner_avg_at_mw_z + asq_total_mw_z, data = dem_plus_mw_clin_cent)
summary(gvlma(m1)) # acceptable
summary(m1)
vif(m1)

# CI
dem_plus_mw_clin_cent <-
  dem_plus_mw_clin_cent %>%
  drop_na(asq_total_mw_z)
m1 <- lm(mw_dailysleepsat_z ~ admn_lecn_mw_z + pdmn_lecn_mw_z + tanner_avg_at_mw_z + asq_total_mw_z, data = dem_plus_mw_clin_cent)
dmnlecnci = Boot(m1, coef,
method='case', R=1000)
dmnlecnciconfint <- confint(dmnlecnci, type='norm')

# BF
dmnlecnbf <- regressionBF(
  formula = mw_dailysleepsat_z ~ admn_lecn_mw_z + pdmn_lecn_mw_z + tanner_avg_at_mw_z + asq_total_mw_z, data = dem_plus_mw_clin_cent
)


m2 <- lm(mw_dailysleepsat_z  ~ admn_recn_mw_z + pdmn_recn_mw_z + tanner_avg_at_mw_z + asq_total_mw_z, data = dem_plus_mw_clin_cent)
summary(gvlma(m2)) # acceptable
summary(m2)

vif(m2)
# CI
dmnrecnci = Boot(m2, coef,
method='case', R=1000)
dmnrecnciconfint <- confint(dmnrecnci, type='norm')

# BF
dmnrecnbf <- regressionBF(
  formula = mw_dailysleepsat_z  ~ admn_recn_mw_z + pdmn_recn_mw_z + tanner_avg_at_mw_z, data = dem_plus_mw_clin_cent
)
```

# associations between within-DMN connectivity and sleep
- given we didn't preregister for parts of the  DMN, we will test the anteiror and posterior in one model
```{r}
m3 <- lm(mw_dailysleepsat_z ~ aDMN_Z_36R_z + pDMN_Z_36R_z + tanner_avg_at_mw_z + asq_total_mw_z, data = dem_plus_mw_clin_cent)
summary(gvlma(m3))
summary(m3)

# CI
apdmnci = Boot(m3, coef,
method='case', R=1000)
apdmnciciconfint <- confint(apdmnci, type='norm')


# BF
apdmnbf <- regressionBF(
  formula = mw_dailysleepsat_z  ~ aDMN_Z_36R_z + pDMN_Z_36R_z + tanner_avg_at_mw_z, data = dem_plus_mw_clin_cent
)

```
> We did not find evidence for effects of inter-network connectivity between each of the lateralized ECNs and the aDMN and pDMN or within-network connectivity in relation to sleep satisfaction. However, pubertal stage and perceived stress were associated with sleep satisfaction in all regressions.

```{r}
library(stargazer)
table_brain <- 
  stargazer(
    m1,m2,m3, 
    type = "text", style = "default",
    title = "Table XX. Associations between FC, Tanner, and Perceived Stress with Sleep Satisfaction",
    dep.var.labels = c("Sleep Satisfaction"),
    covariate.labels = c("Intercept", "aDMN-lECN", "pDMN-lECN", "aDMN-rECN", "pDMN-rECN", "aDMN", "pDMN", "Tanner Stage", "Perceived Stress"), 
    summary = TRUE, 
    out.header = TRUE,
    single.row = TRUE, 
    ci=TRUE, 
    ci.level = .95, 
    intercept.bottom = FALSE, 
    intercept.top = FALSE, 
    digits = 2
    )
write_lines(table_brain, "table_brain.tsv")
```


--------------------------------------------------------------------------------
# sleep duration (supplementary)

```{r}

sleepout = boxplot(dem_plus_mw_clin_netcorr$dailysleep_hrs_mean)$out 
# 3 outliers
which(dem_plus_mw_clin_netcorr$dailysleep_hrs_mean %in% sleepout)
dem_plus_mw_clin_netcorr[which(dem_plus_mw_clin_netcorr$dailysleep_hrs_mean %in% sleepout),]
dem_plus_mw_clin_netcorr_rem <- 
  dem_plus_mw_clin_netcorr %>%
  filter(
    ELS_ID != "5",
    ELS_ID != "20",
    ELS_ID != "193"
  )
```

## centering
I standardized tanner stage, age at tanner assessment, and bmi within each sex
```{r}
dem_plus_mw_clin_cent <-
  dem_plus_mw_clin_netcorr_rem %>%
  mutate(
    Sex = factor(Sex),
    mw_dailysleepsat_z = scale(dailysleep_sat_mean),
    mw_dailysleephrs_z = scale(dailysleep_hrs_mean),
    COVID_AHC_ACT = factor(COVID_AHC_ACT),
    aDMN_Z_36R_z = scale(aDMN_Z_36R),
    pDMN_Z_36R_z = scale(pDMN_Z_36R),
    admn_lecn_mw_z = scale(admn_lecn_mw),
    pdmn_lecn_mw_z = scale(pdmn_lecn_mw),
    admn_recn_mw_z = scale(admn_recn_mw),
    pdmn_recn_mw_z = scale(pdmn_recn_mw),
    Noise_brainstem_Z_36R_z = scale(Noise_brainstem_Z_36R)    
    ) %>%
  group_by(Sex) %>%
  mutate(
    tanner_avg_at_mw_z = scale(tanner_average_for_mw), # standardize within each sex
    tanner_adrenal_mw_z = scale(tanner_adrenal_mw),
    tanner_gonadal_mw_z = scale(tanner_gonadal_mw),
    bmi_at_mw_z = scale(BMI_MW),
    age_at_tanner_at_mw_z = scale(tanner_age_for_mw)
  ) %>%
  ungroup()
```


# test of potential covariates
```{r}
# corr btwn sleep sat and interval between mw and scan
cor.test(dem_plus_mw_clin_cent$mw_dailysleephrs_z, dem_plus_mw_clin_cent$interval_btwn_mw_and_scan) # no corr, r = -0.06742619, p = 0.5709

# corr btwn sleep sat and mean FD
cor.test(dem_plus_mw_clin_cent$mw_dailysleephrs_z, dem_plus_mw_clin_cent$meanFD) # no corr, r = 0.1927567, p = 0.1023

# corr btwn sleep sat and age
cor.test(dem_plus_mw_clin_cent$mw_dailysleephrs_z, dem_plus_mw_clin_cent$age_at_tanner_at_mw_z)  # no corr, r = -0.0570573 , p = 0.6316 (as seen in our puberty analysis)

# corr btwn sleep sat and tanner
cor.test(dem_plus_mw_clin_cent$mw_dailysleephrs_z, dem_plus_mw_clin_cent$tanner_avg_at_mw_z) # yes, corr, r = -0.2310467, p = 0.04922 (as seen in our puberty analysis)

# association btwn sleep sat and sex
summary(lm(mw_dailysleephrs_z ~ Sex, data = dem_plus_mw_clin_cent)) # no association, B = -0.3073, p = 0.199

# corr btwn sleep sat and dperession severity
cor.test(dem_plus_mw_clin_cent$mw_dailysleephrs_z, dem_plus_mw_clin_cent$cdi_total_with_mw)  # no corr, r = -0.2914324, p = 0.0159

# corr btwn sleep sat and SI severity
cor.test(dem_plus_mw_clin_cent$mw_dailysleephrs_z, dem_plus_mw_clin_cent$siq_total_with_mw)  # no corr, r = -0.1941069, p = 0.1155)

# corr btwn sleep sat and perceived stress
cor.test(dem_plus_mw_clin_cent$mw_dailysleephrs_z, dem_plus_mw_clin_cent$asq_total_mw)  # yes corr, r = -0.2782732, p = 0.02722
```

# associations between inter-network connectivity and sleep
```{r}
options(scipen = 999)

m1 <- lm(mw_dailysleephrs_z ~ admn_lecn_mw_z + pdmn_lecn_mw_z + cdi_total_with_mw + siq_total_with_mw, data = dem_plus_mw_clin_cent)
summary(gvlma(m1)) # acceptable
summary(m1)

m2 <- lm(mw_dailysleephrs_z  ~ admn_recn_mw_z + pdmn_recn_mw_z + cdi_total_with_mw + siq_total_with_mw, data = dem_plus_mw_clin_cent)
summary(gvlma(m2)) # acceptable
summary(m2)

```

# associations between within-DMN connectivity and sleep
- given we didn't preregister for parts of the  DMN, we will test the anteiror and posterior in one model
```{r}
m3 <- lm(mw_dailysleephrs_z ~ aDMN_Z_36R_z + pDMN_Z_36R_z + cdi_total_with_mw + siq_total_with_mw, data = dem_plus_mw_clin_cent)
summary(gvlma(m3))
summary(m3)
```
