---
title: "clust_stand_err_final"
author: "Jackie"
date: "7/21/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---


# libraries
```{r,message=FALSE}
library(tidyverse)
library(lmtest)
library(lubridate)
library(pmdplyr)
library(sandwich)
library(performance)
library(sjPlot)
library(sjstats)
```

## functions
```{r}
getmode <- function(v, na.rm=TRUE) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
# function from ben bolker
mylag <- function(x,lag) {
  c(rep(NA,lag),head(x,-lag))
}
```

# Read in data

## Aggregating 
In these types of data sampling situations (different frequency of assessment) – the typical solution for examination of coupling seems to be to aggregate to a common time-sampling frame (e.g., day-level) and locate coupling at that timescale.

Bc my design was such that participants report their evening affect and next day sleep satisfactionj
## sleep: EMA

```{r, message=FALSE}
mw_sleep_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/0_MW_Act_Demo_Descriptives/MW_daily_sleep_data_longform.csv"
mw_sleep <- read_csv(mw_sleep_fp) %>%
  mutate(
    ELS_ID = factor(ELS_ID),
    day = factor(day),
    wday = factor(wday),
    dayorder = factor(dayorder),
    week = factor(week),
    trigger_tod = "Morning",
    trigger_tod = factor(trigger_tod)
    ) %>%
  dplyr::select(
    ELS_ID,
    triggerdate,
    day,
    dayorder,
    sleep_trigger_time,
    trigger_tod,
    wday,
    week,
    everything(),
    -Session
  ) %>%
  dplyr::rename(., triggertime = sleep_trigger_time) %>%
  mutate(
    day = factor(day)
  ) %>%
  mutate(
    triggerdate = ymd(triggerdate),
    day = day(triggerdate)
  ) %>%
  group_by(ELS_ID) %>% 
  mutate(
    dayorder = order(triggerdate)
    ) %>%
  dplyr::select(
    ELS_ID,
    triggerdate,
    day,
    dayorder,
    trigger_tod,
    starts_with("DailySleep")
  ) %>%
  mutate(
    trigger_tod_order =
      ifelse(
        trigger_tod == "Morning",
        "1",
        NA
        )
    ) %>%
  relocate(trigger_tod_order, .after = trigger_tod) %>%
  mutate(
    trigger_tod_order = as.numeric(trigger_tod_order)
  ) %>%
  group_by(ELS_ID) %>%
  mutate(
    dayorder = as.numeric(dayorder),
    respnumber = dayorder - min(dayorder) + 1
  ) %>%
  filter(dayorder < 15) %>%
  relocate(respnumber , .after = dayorder) %>%
  mutate(
    day_id =  time_variable(triggerdate, .method = "day")
    ) %>%
  relocate(day_id, .after = day) %>%
  filter(day_id < 15) %>% 
  dplyr::select(-c(day, dayorder)) # already aggregated by day
  
mw_sleep_select <-
  mw_sleep %>%
  dplyr::select(-c(day_id, respnumber, DailySleep_hrs))
```

## affect: EMA

```{r, message=FALSE}
mw_emo_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/2_study2/MW_daily_emo_data_longform.csv"
mw_emo <- read_csv(mw_emo_fp) %>%
  dplyr::select(
    ELS_ID,
    mw_start_date,
    emo_trigger_time,
    emo_trigger_recode,
    triggerdate,
    day,
    wday,
    dayorder,
    day_id,
    neg_total,
    pos_total,
    data_num_obs,
    n_per_tod,
    week
  ) %>%
  rename(
    respnumber = dayorder
  )

mw_emo_filt <-
  mw_emo %>%
  mutate(
    trigger_tod_order =
      ifelse(
        emo_trigger_recode == "Morning",
        "1",
        NA
        ),
    trigger_tod_order =
      ifelse(
        emo_trigger_recode == "Afternoon",
        "2",
        trigger_tod_order
        ),
    trigger_tod_order =
      ifelse(
        emo_trigger_recode == "Evening",
        "3",
        trigger_tod_order
        )
    ) %>%
  relocate(trigger_tod_order, .after = emo_trigger_recode) %>%
  mutate(
    trigger_tod_order = as.numeric(trigger_tod_order),
    iddrop =
      ifelse(
        ELS_ID == "199" &
          str_detect(mw_start_date, "2021-02"),
        "drop",
        NA
      )
  ) %>%
  filter(is.na(iddrop)) %>%
  dplyr::select(-iddrop) %>%
  mutate(
    ELS_ID = factor(ELS_ID)
  )

## reading in RR
emorrfp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/mw_daily_emo_rr.csv" 
emorr <- 
  read_csv(emorrfp) %>%
  dplyr::select(
    ELS_ID,
    Enrolled,
    Daily_Emotion_Responses,
    emo_resp_rate
  ) %>%
  mutate(
    dup = duplicated(ELS_ID), # 120 (april 2019), 201 (July 2020), 199 (april 2021 - the one with 7 responses)
    drop =
      ifelse(
        ELS_ID == "120" &
          Enrolled == "09-07-2021 10:35:28",
        "drop",
        NA
        ),
    drop =
      ifelse(
        ELS_ID == "201" &
          Enrolled == "Not Enrolled",
        "drop",
        drop
      ),
    drop = 
      ifelse(
        ELS_ID == "199" &
          Daily_Emotion_Responses == "31",
        "drop",
        drop
      )
  ) %>%
  filter(is.na(drop)) %>%
  dplyr::select(-drop, -dup) %>%
  mutate(
    ELS_ID = as.numeric(ELS_ID),
    ELS_ID = factor(ELS_ID)
  )
  
# merging RR with the daily emo df
mw_emo_filt_merge <-
  left_join(
    mw_emo_filt,
    emorr,
    df = "ELS_ID"
  ) %>%
  rename(trigger_tod = emo_trigger_recode) %>%
  mutate(trigger_tod = factor(trigger_tod))
```

```{r, message=FALSE}
mw_icafp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/3_study3/mw_36R_clean_sens.csv" 
mw_ica <- 
  read_csv(mw_icafp) %>%
  mutate(ELS_ID = factor(ELS_ID)) 
```

# Merging

```{r}
# just the EMA stuff
mw_merge <-
  left_join(
    mw_emo_filt_merge,
    mw_sleep_select,
    by = c("ELS_ID", "triggerdate", "trigger_tod", "trigger_tod_order")
  ) %>%
  mutate(
    triggerdate = ymd(triggerdate),
    day = day(triggerdate)
  ) %>%
  group_by(ELS_ID) %>% 
  mutate(
    day_id =  time_variable(triggerdate, .method = "day"),    
    dayorder = order(triggerdate),
    dayorder = as.numeric(dayorder),
    respnumber = dayorder - min(dayorder) + 0,
    day_id = day_id - min(day_id) + 0,
    trigger_tod_order = trigger_tod_order - min(trigger_tod_order) + 0
    ) %>%
  dplyr::select(
    ELS_ID, triggerdate, day_id, respnumber, trigger_tod, trigger_tod_order, DailySleep_satisfaction, DailySleep_hrs_rec, neg_total, pos_total, emo_resp_rate 
  )

mw_merge %>%
  distinct(ELS_ID) # 131
```

### merging with brain data in case I decide to look at this based on my other brain findings

```{r}
mw_merge <-
  left_join(
    mw_merge,
    mw_ica,
    by = "ELS_ID"
  )
```

## MW EMO Start Date
```{r,message=FALSE}
emo_rr_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/mw_daily_emo_rr.csv"
emo_rr <- 
  read_csv(emo_rr_fp) %>%
  dplyr::select(
    ELS_ID, Identifier, Enrolled, emo_resp_rate
  ) %>%
  rename(User_ID = Identifier) %>%
  mutate(
    ELS_ID = as.numeric(ELS_ID),
    ELS_ID = factor(ELS_ID),
    dup = duplicated(User_ID)
  )  %>% # keeping ELS_ID 201 and 199 with april enrolled date
mutate(
    drop = 
      ifelse(
        ELS_ID == "201" &
          Enrolled == "Not Enrolled",
        "drop",
        NA
        )
  ) %>%
  filter(is.na(drop)) %>%
  dplyr::select(ELS_ID, emo_resp_rate)
```
# Reading in pubertal and demo data

> the suffix "_mw" or "_act" indicates whether that Tanner assessemnt was administered closer to the actigraphy or EMA data. 154 participants had usable actigraphy or EMA data; however, 6 participants were excluded based on being in early puberty (Tanner stage < 3), bringing our total to 148 participants. 136 had usable actigraphy data and 111 had usable EMA data via Metricwire data. 54% of our sample (N=80) completed actigraphy and EMA components simultaneously (overlapping at least 7 days). This was not by design, but because of participant scheduling/contacting issues or bc of participant issues, the other half of the participants either completed both components several months apart from each other, _or_ they completed only one of the componets (either EMA _or_ actigraphy due to various reasons (e.g., didn't want to wear a watch, didn't want the app on their phones, data from one component wasn't usable, etc...)).
To include as many participants as possible in our analyses (to increase generalizability and statistical power), we conducted the EMA and actigraphy analyses separately.

```{r, message=FALSE}
# full AHC sample with pubertal and demographic info
pub_demo_act_mw_comb_fp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/0_MW_Act_Demo_Descriptives/dem_and_tanner_mw_act_comb_merge.csv"
pub_demo <-
  read_csv(pub_demo_act_mw_comb_fp) %>%
  mutate(ELS_ID = factor(ELS_ID))%>%
  dplyr::select(
    ELS_ID,
    SIMUL_AHC,
    tanner_date_for_act,
    tanner_date_for_mw,
    tanner_average_for_act,
    tanner_average_for_mw,
    tanner_adrenal_act,
    tanner_adrenal_mw,
    tanner_gonadal_act,
    tanner_gonadal_mw,
    tanner_age_for_act,
    tanner_age_for_mw,
    COVID_AHC_MW,
    COVID_AHC_ACT,
    Sex,
    BMI_MW,
    BMI_ACT,
    Household_Income_MW,
    Household_Income_ACT,
    SIMUL_AHC,
    cmep_total_mw,
    cmep_total_act,
    sumsev_type
  )

hormonedatafp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/1_study1/pub_mw_hormones_clean.csv"
hormonedata <-
  read_csv(hormonedatafp) %>%
  dplyr::select(
    ELS_ID,
    ends_with("_clean")
  ) %>%
  mutate(ELS_ID = factor(ELS_ID))

asqfp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/2_study2/asqmw.csv"
asq <-
  read_csv(asqfp) %>%
  mutate(ELS_ID = factor(ELS_ID))
```

## merging with EMA data
```{r}
mw_merge_order_dem_all <-
  left_join(
    mw_merge,
    pub_demo,
    by = "ELS_ID"
  )

mw_merge_order_dem_all <-
  left_join(
    mw_merge_order_dem_all,
    hormonedata,
    by = "ELS_ID"
  )

mw_merge_order_dem_all <-
  left_join(
    mw_merge_order_dem_all,
    asq,
    by = "ELS_ID"
  ) %>%
  drop_na(tanner_average_for_mw)
```

# Summary

```{r}
mw_merge_sum_ema <-
  mw_merge_order_dem_all %>%
  distinct(ELS_ID)
# 106 IDs have EMA data (sleep + affect); 

mw_merge_order_dem_all%>%
  group_by(ELS_ID) %>%
  summarize(
    daymax = length(day_id)
  )

mw_merge_order_dem_all2 <-
  mw_merge_order_dem_all %>%
  group_by(ELS_ID) %>%
  mutate(
    daymax = length(day_id),
    toofewdays =
      ifelse(
        daymax < 3, # dropping fewer than 3 days of observations
        "drop",
        NA
        )
    ) %>%
  filter(
    is.na(toofewdays)
  ) %>%
  dplyr::select(-daymax, -toofewdays)

mw_merge_order_dem_all2 %>%
  group_by(ELS_ID) %>%
  summarize(
    daymax = length(day_id)
  )
# 106 participants have 3 or more days of sleep and affect data

```

```{r}
summary(as.integer(table(mw_merge_order_dem_all2$ELS_ID))) # els ids subj
```

```{r}
hist(mw_merge_order_dem_all2$neg_total)
library(moments)

skewness(mw_merge_order_dem_all2$neg_total) # 0.9293557

mw_merge_order_dem_all2 <-
  mw_merge_order_dem_all2 %>%
  mutate(
    neg_total_log = log10(neg_total),
    neg_total_sqrt = sqrt(neg_total)
  )
skewness(mw_merge_order_dem_all2$neg_total_log) # -1.301284 ooo terrible
skewness(mw_merge_order_dem_all2$neg_total_sqrt) # -0.04143427 much better


summary(mw_merge_order_dem_all2$neg_total_sqrt)
mean(mw_merge_order_dem_all2$neg_total_sqrt)
sd(mw_merge_order_dem_all2$neg_total_sqrt)
```


```{r}
neg_total_sqrt_hist <- 
  mw_merge_order_dem_all2 %>%
  ggplot(
    aes(x = neg_total_sqrt)
  ) +
  geom_histogram(color = "black", alpha = .5) +
  theme_classic() +
  labs(
    x = "Daily Negative Affect",
    title = "Distribution of Daily Negative Affect"
  ) + 
  theme(
    axis.text = element_text(size = 10, angle = 30, hjust = 1),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 10)
  )
neg_total_sqrt_hist


pos_total_sqrt_hist <- 
  mw_merge_order_dem_all2 %>%
  ggplot(
    aes(x = pos_total)
  ) +
  geom_histogram(color = "black", alpha = .5) +
  theme_classic() +
  labs(
    x = "Daily Positive Affect",
    title = "Distribution of Daily Positive Affect"
  ) + 
  theme(
    axis.text = element_text(size = 10, angle = 30, hjust = 1),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 10)
  )
pos_total_sqrt_hist

library(gridExtra)
library(cowplot)
pana_dist <- 
  cowplot::plot_grid(
    neg_total_sqrt_hist, 
    pos_total_sqrt_hist,
    labels = c('A', 'B'), 
    label_size = 10
    )
ggsave("pana_dist.png", pana_dist, width = 7, height = 3.5)
```
## how many clusters do we have?
```{r}
sumclust <-
  mw_merge_order_dem_all2 %>%
  distinct(ELS_ID)
```

# Option 1: the sandwich package
The `sandwich` package implements several methods for robust covariance estimators, including clustered SEs. Details are explained in __Zeileis et al. 2020__. The accompanying `lmtest` package provides functions for coefficient tests that take into account the calculated robust covariance estimates.

As explained initially, the parameter estimates from our model are fine despite the clustered structure of our data. But the SEs are likely biased downward and need to be corrected.

1) We need to use `coeftest` from the `lmtest` package;
2) we need to pass it our model and either a function to calculate the covariance matrix or an already estimated covariance matrix to the `vcov` parameter;
3) we need to specify a cluster variable in the cluster parameter.

The `sandwich` package provides several functions for estimating robust covariance matrices. We need `vcovCL` for clustered covariance estimation and will pass this function as vcov parameter. Furthermore, we cluster by subject ID, so the cluster variable is ELS_ID
# Clustered Robust Standard Errors (CR-SE)

From "On the Unnecessary Ubiquity of Hierarchical Linear Modeling" by McNeish et al., 2016
"CR-SEs alter the calculation of the regression coefficient variance to accommodate assumption violations (i.e., homogeneity of variance and independence).
Conceptually, with CR-SEs, first a standard single-level model (e.g., with OLS or maximum likelihood as shown in Equation 3) is used to estimate the regression coefficients (e.g., ). In the presence of clustering, bias in the standard error estimates are the primary concern, so the regression coefficients with CR-SEs will be identical to what would be obtained if the clustering was completely ignored."

## showing clusters of ELS_IDs in relation to neg affect
```{r}
mw_merge_cent <-
  mw_merge_order_dem_all2 %>%
  dplyr::select(
    -c(ends_with("_act"), ends_with("_ACT"))
  ) %>%
  mutate(
    Sex = factor(Sex),
    COVID_AHC_MW = factor(COVID_AHC_MW),
    sumsev_type_z = scale(sumsev_type),
    dhea_clean_z = scale(dhea_clean),
    cort_clean_z = scale(cort_clean),
    test_clean_z = scale(test_clean),
    cmep_total_mw_z = scale(cmep_total_mw),
    asq_total_mw_z = scale(asq_total_mw)
    ) %>%
  group_by(Sex) %>%
  mutate(
    tanner_average_for_mw_z = scale(tanner_average_for_mw), # standardizing tanner average within each sex
    tanner_adrenal_mw_z = scale(tanner_adrenal_mw), # standardizing tanner adrenal within each sex
    tanner_gonadal_mw_z = scale(tanner_gonadal_mw), # standardizing tnaner gonadal within each sex
    bmi_at_mw_z = scale(BMI_MW), # standardizing bmi within each sex
    tanner_age_for_mw_z = scale(tanner_age_for_mw) # standardizing age within each sex
  ) %>%
  ungroup() 

library(lmerTest)
mw_merge_cent_randids <-
  mw_merge_cent[1:234,]
mw_merge_cent_randids %>%
  ggplot(
    aes(
      x = day_id,
      y = neg_total_sqrt,
      color = ELS_ID
    )
  ) +
  geom_point()+
  facet_wrap(~ELS_ID) +
  theme_classic()

library(misty)
multilevel.icc(mw_merge_cent$neg_total_sqrt, group = mw_merge_cent$ELS_ID, type = 1, method = "lme4") # 0.6100043
mod <- lmer(neg_total_sqrt ~ 1 + (1|ELS_ID), data = mw_merge_cent)
icc(mod)
multilevel.icc(mw_merge_cent$pos_total, group = mw_merge_cent$ELS_ID, type = 1, method = "lme4") # 0.4910874
multilevel.icc(mw_merge_cent$DailySleep_satisfaction, group = mw_merge_cent$ELS_ID, type = 1, method = "lme4") # 0.4152467
```

# Hypothesis 1: Higher negative affect and lower positive affect and will be associated with lower subjective sleep satisfaction, but not with poorer objective sleep efficiency or lower sleep duration
testing for potential covariates
```{r}
c7 <- lm(scale(DailySleep_satisfaction) ~ bmi_at_mw_z, data = mw_merge_cent)
c7_c <- coeftest(c7, vcov = vcovCL, cluster = ~ELS_ID) # nope
c7_c
c7b <- lm(scale(DailySleep_hrs_rec) ~ bmi_at_mw_z, data = mw_merge_cent)
c7_d <- coeftest(c7b, vcov = vcovCL, cluster = ~ELS_ID) # nope
c7_d

c8 <- lm(scale(DailySleep_satisfaction) ~ scale(Household_Income_MW), data = mw_merge_cent)
c8_c <- coeftest(c8, vcov = vcovCL, cluster = ~ELS_ID) # nope
c8_c
c8b <- lm(scale(DailySleep_hrs_rec) ~ scale(Household_Income_MW), data = mw_merge_cent)
c8_d <- coeftest(c8b, vcov = vcovCL, cluster = ~ELS_ID) # nope
c8_d

c9 <- lm(scale(DailySleep_satisfaction) ~ scale(sumsev_type), data = mw_merge_cent)
c9_c <- coeftest(c9, vcov = vcovCL, cluster = ~ELS_ID) # nope
c9_c
c9b <- lm(scale(DailySleep_hrs_rec) ~ scale(sumsev_type), data = mw_merge_cent)
c9_d <- coeftest(c9b, vcov = vcovCL, cluster = ~ELS_ID) # yes
c9_d

c10 <- lm(scale(DailySleep_satisfaction) ~ Sex, data = mw_merge_cent)
c10_c <- coeftest(c10, vcov = vcovCL, cluster = ~ELS_ID) # nope
c10_c
c10b <- lm(scale(DailySleep_hrs_rec) ~ Sex, data = mw_merge_cent)
c10_d <- coeftest(c10b, vcov = vcovCL, cluster = ~ELS_ID) # nope
c10_d


c11 <- lm(scale(DailySleep_satisfaction) ~ scale(asq_total_mw), data = mw_merge_cent)
c11_c <- coeftest(c11, vcov = vcovCL, cluster = ~ELS_ID) # yes B=-0.218765, p=0.003652
c11_c
c11b <- lm(scale(DailySleep_hrs_rec) ~ scale(asq_total_mw), data = mw_merge_cent)
c11_d <- coeftest(c11b, vcov = vcovCL, cluster = ~ELS_ID) # yes B=-0.218765, p=0.003652
c11_d
```

### making sure my pubertal stage and sleep satisfaction association holds using this type of model (adjusting for clustered IDs)
```{r}
m2 <- lm(scale(DailySleep_satisfaction) ~  tanner_average_for_mw_z + tanner_age_for_mw_z, data = mw_merge_cent)
# also holds controlling for asq
m2cov <- lm(scale(DailySleep_satisfaction) ~  scale(asq_total_mw) + tanner_average_for_mw_z + tanner_age_for_mw_z, data = mw_merge_cent)
summary(m2cov)
car::vif(m2)
m2_c <- coeftest(m2, vcov = vcovCL, cluster = ~ELS_ID)
m2_c
confint(m2_c)
m2_cov <- coeftest(m2cov, vcov = vcovCL, cluster = ~ELS_ID)
m2_cov
confint(m2_cov)

tanint2 <- summary(m2)$coefficients[1,1] # fine to extract from the OLS bc reg coefs don't change (just the SEs)
tanslp2 <- summary(m2)$coefficients[2,1]

mw_merge_cent %>%
  ggplot(
    aes(
      x = tanner_average_for_mw_z,
      y = scale(DailySleep_satisfaction)
    )
  ) +
  geom_point(size = 2, alpha = .5) +
  geom_abline(
    intercept = tanint2, 
    slope = tanslp2,
    size=2
    ) +
  theme_classic() +
  labs(
    x = "Tanner Score",
    y = "Sleep Satisfaction",
    title = "Association between Pubertal Stage and Sleep Satisfaction"
  ) + 
  theme(
    axis.text = element_text(size = 12, angle = 30, hjust = 1),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 12)
  ) 

ggsave("tan_and_sleep_crse.png", width = 6, height =6)


# MLM
level2modds<- lmer(scale(DailySleep_satisfaction)  ~  tanner_average_for_mw_z + scale(asq_total_mw) + tanner_age_for_mw_z + day_id + (1| ELS_ID), na.action = na.exclude, data = mw_merge_cent)
tab_model(level2modds, p.val = "kr") 
# aligns with cr-se
```

## testing if postive and negative affect are associated with sleep sat controlling for perceived stress (and then also tanner)
```{r}
m3 <- lm(scale(DailySleep_satisfaction) ~ scale(asq_total_mw) + scale(neg_total_sqrt) + scale(pos_total), data = mw_merge_cent)
summary(m3)
car::vif(m3)
m3c <- coeftest(m3, vcov = vcovCL, cluster = ~ELS_ID)
confint(m3c)


m3cov <- lm(scale(DailySleep_satisfaction) ~ scale(asq_total_mw) + scale(neg_total_sqrt) + scale(pos_total) +  tanner_average_for_mw_z +  tanner_age_for_mw_z, data = mw_merge_cent)
summary(m3cov)
car::vif(m3cov)
m3cov_c <- coeftest(m3cov, vcov = vcovCL, cluster = ~ELS_ID)
confint(m3cov_c)

negint3 <- summary(m3)$coefficients[1,1] # fine to extract from the OLS bc reg coefs don't change (just the SEs)
negslp3 <- summary(m3)$coefficients[3,1]
na_sleep <- mw_merge_cent %>%
  ggplot(
    aes(
      x = scale(neg_total_sqrt),
      y = scale(DailySleep_satisfaction)
    )
  ) +
  geom_point(size = 2, alpha = .5, color = "black") +
  geom_abline(
    intercept = negint3, 
    slope = negslp3,
    size=2,
    color = "black"
    ) +
  theme_classic() +
  labs(
    x = "Negative Affect",
    y = "Sleep Satisfaction"
  ) + 
  theme(
    axis.text = element_text(size = 10, angle = 30, hjust = 1),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 12)
  )

posint3 <- summary(m3)$coefficients[1,1] # fine to extract from the OLS bc reg coefs don't change (just the SEs)
posslp3 <- summary(m3)$coefficients[4,1]
pa_sleep <- mw_merge_cent %>%
  ggplot(
    aes(
      x = scale(pos_total),
      y = scale(DailySleep_satisfaction)
    )
  ) +
  geom_point(size = 2, alpha = .5, color = "black") +
  geom_abline(
    intercept = posint3, 
    slope = posslp3,
    size=2,
    color = "black"
    ) +
  theme_classic() +
  labs(
    x = "Positive Affect",
    y = "Sleep Satisfaction"
  ) + 
  theme(
    axis.text = element_text(size = 10, angle = 30, hjust = 1),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 12)
  )
library(gridExtra)
library(cowplot)
mwsleepaffect <- 
  cowplot::plot_grid(
    na_sleep, 
    pa_sleep, 
    labels = c('A', 'B'), 
    label_size = 10
    )
ggsave("sleep_NA_PA_crse.png", mwsleepaffect, width = 9, height = 5)
# MLM
level3modds<- lmer(scale(DailySleep_satisfaction)  ~  tanner_average_for_mw_z + scale(asq_total_mw) + scale(neg_total_sqrt) + scale(pos_total) + day_id + (1| ELS_ID), na.action = na.exclude, data = mw_merge_cent)
tab_model(level3modds, p.val = "kr") 
# aligns with cr-se
```

## testing if postive and negative affect are associated with sleep duration controlling for perceived stress (and then also tanner)
```{r}

m4 <- lm(scale(DailySleep_hrs_rec) ~ scale(sumsev_type) + scale(neg_total_sqrt) + scale(pos_total), data = mw_merge_cent)
summary(m4)
car::vif(m4)
m4c <- coeftest(m4, vcov = vcovCL, cluster = ~ELS_ID)
confint(m4c)
```


# Hypothesis 2:  Higher Tanner stage is associated with higher negative affect and lower positive affect
testing for potential covariates
```{r}
c1 <- lm(scale(neg_total_sqrt) ~ bmi_at_mw_z, data = mw_merge_cent)
c1_c <- coeftest(c1, vcov = vcovCL, cluster = ~ELS_ID) # nope
c1_c

c2 <- lm(scale(neg_total_sqrt) ~ scale(Household_Income_MW), data = mw_merge_cent)
c2_c <- coeftest(c2, vcov = vcovCL, cluster = ~ELS_ID) # nope
c2_c

c3 <- lm(scale(neg_total_sqrt) ~ scale(sumsev_type), data = mw_merge_cent)
c3_c <- coeftest(c3, vcov = vcovCL, cluster = ~ELS_ID) # nope
c3_c

mw_merge_cent <-
  mw_merge_cent %>%
  mutate(
    Sex_rec =
      recode_factor(Sex, "1" = "Male", "2" = "Female")
  )
c4 <- lm(scale(neg_total_sqrt) ~ Sex_rec, data = mw_merge_cent)
c4_c <- coeftest(c4, vcov = vcovCL, cluster = ~ELS_ID) # yes (males compared to females have lower neg affect)
c4_c
mw_merge_cent %>%
  filter(Sex_rec == "Female") %>%
  summarise(
    neg_total_mean = mean(neg_total),
    neg_total_sd = sd(neg_total)
  )
mw_merge_cent %>%
  filter(Sex_rec == "Male") %>%
  summarise(
    neg_total_mean = mean(neg_total),
    neg_total_sd = sd(neg_total)
  )

c5 <- lm(scale(neg_total_sqrt) ~ COVID_AHC_MW, data = mw_merge_cent)
c5_c <- coeftest(c5, vcov = vcovCL, cluster = ~ELS_ID) # no
c5_c

c6 <- lm(scale(neg_total_sqrt) ~ scale(asq_total_mw), data = mw_merge_cent)
c6_c <- coeftest(c6, vcov = vcovCL, cluster = ~ELS_ID) # yes
c6_c

c7 <- lm(scale(neg_total_sqrt) ~ scale(emo_resp_rate), data = mw_merge_cent)
c7_c <- coeftest(c7, vcov = vcovCL, cluster = ~ELS_ID) # no
c7_c
```

```{r}
# testing whether tanner_average on its own is associated with NA
m1 <- lm(scale(neg_total_sqrt) ~ tanner_average_for_mw_z + tanner_age_for_mw_z, data = mw_merge_cent)
summary(m1)
car::vif(m1) # no multicollinearity
m1_c <- coeftest(m1, vcov = vcovCL, cluster = ~ELS_ID)
m1_c
confint(m1_c)

## adding covariates of Sex and Adolescent perceived stress
m1_cov <- lm(scale(neg_total_sqrt) ~ tanner_average_for_mw_z + tanner_age_for_mw_z +Sex + scale(asq_total_mw), data = mw_merge_cent)
summary(m1_cov) # effects retained
car::vif(m1_cov) # no multicollinearity
m1_cov_c <- coeftest(m1_cov, vcov = vcovCL, cluster = ~ELS_ID)
m1_cov_c
confint(m1_cov_c)

tanint <- summary(m1)$coefficients[1,1] # fine to extract from the OLS bc reg coefs don't change (just the SEs)
tanslp <- summary(m1)$coefficients[2,1]

mw_merge_cent %>%
  ggplot(
    aes(
      x = tanner_average_for_mw_z,
      y = scale(neg_total_sqrt)
    )
  ) +
  geom_jitter(size = 2, alpha = .3) +
  geom_abline(
    intercept = tanint,
    slope = tanslp,
    size = 2
    )+
  theme_classic() +
  labs(
    x = "Tanner Average",
    y = "Negative Affect",
    title = "Association between Pubertal Stage and Negative Affect"
  ) + 
  theme(
    axis.text = element_text(size = 10, angle = 30, hjust = 1),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 10)
  )
ggsave("tan.negaffect_crse.png", width = 6, height = 6)

# MLM
level2modna<- lmer(neg_total_sqrt ~  tanner_average_for_mw_z + tanner_age_for_mw_z + scale(asq_total_mw) + Sex + day_id + (1| ELS_ID), na.action = na.exclude, data = mw_merge_cent)
tab_model(level2modna, p.val = "kr") 
# aligns with cr-se
```

```{r}
# testing whether tanner_average on its own is associated with pa
c1 <- lm(scale(pos_total) ~ bmi_at_mw_z, data = mw_merge_cent)
c1_c <- coeftest(c1, vcov = vcovCL, cluster = ~ELS_ID) # nope
c1_c
c2 <- lm(scale(pos_total) ~ scale(Household_Income_MW), data = mw_merge_cent)
c2_c <- coeftest(c2, vcov = vcovCL, cluster = ~ELS_ID) # nope
c2_c
c3 <- lm(scale(pos_total) ~ scale(sumsev_type), data = mw_merge_cent)
c3_c <- coeftest(c3, vcov = vcovCL, cluster = ~ELS_ID) # nope
c3_c
c4 <- lm(scale(pos_total) ~ Sex_rec, data = mw_merge_cent)
c4_c <- coeftest(c4, vcov = vcovCL, cluster = ~ELS_ID) # no
c4_c
mw_merge_cent %>%
  filter(Sex_rec == "Female") %>%
  summarise(
    pamean = mean(pos_total),
    pasd = sd(pos_total)
  )
mw_merge_cent %>%
  filter(Sex_rec == "Male") %>%
  summarise(
    pamean = mean(pos_total),
    pasd = sd(pos_total)
  )


c5 <- lm(scale(pos_total) ~ COVID_AHC_MW, data = mw_merge_cent)
c5_c <- coeftest(c5, vcov = vcovCL, cluster = ~ELS_ID) # no
c5_c
c6 <- lm(scale(pos_total) ~ scale(asq_total_mw), data = mw_merge_cent)
c6_c <- coeftest(c6, vcov = vcovCL, cluster = ~ELS_ID) # yes
c6_c
c7 <- lm(scale(pos_total) ~ scale(emo_resp_rate), data = mw_merge_cent)
c7_c <- coeftest(c7, vcov = vcovCL, cluster = ~ELS_ID) # no
c7_c

m1pa <- lm(scale(pos_total) ~ tanner_average_for_mw_z + tanner_age_for_mw_z, data = mw_merge_cent)
summary(m1pa)
m1pa_c <- coeftest(m1pa, vcov = vcovCL, cluster = ~ELS_ID)
m1pa_c
confint(m1pa_c)

m1pacov <- lm(scale(pos_total) ~ tanner_average_for_mw_z + tanner_age_for_mw_z + scale(asq_total_mw), data = mw_merge_cent)
summary(m1pacov)
car::vif(m1pacov) # no multicollinearity
m1pa_c_cov <- coeftest(m1pacov, vcov = vcovCL, cluster = ~ELS_ID)
m1pa_c_cov
confint(m1pa_c_cov)

# MLM
level2modpa<- lmer(scale(pos_total)  ~  tanner_average_for_mw_z + tanner_age_for_mw_z + scale(asq_total_mw) + day_id + (1| ELS_ID), na.action = na.exclude, data = mw_merge_cent)
tab_model(level2modpa, p.val = "kr") 
# aligns with cr-se
```


## is objective sleep associated with affect

### reading in data

```{r,message=FALSE}
actfp <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/0_MW_Act_Demo_Descriptives/SimultaneousAHC/act_mw_simultaneous.csv"
act <-
  read_csv(actfp) %>%
  mutate(
    ELS_ID = factor(ELS_ID)
  )
demoact <- "~/Box/Mooddata_Coordinating/1_Lab_Coordinating/Users/JackieSchwartz/Dissertation/0_MW_Act_Demo_Descriptives/SimultaneousAHC/dem_and_tanner_mw_act_simultaneous.csv"
demoact<-
  read_csv(demoact) %>%
  mutate(
    ELS_ID = factor(ELS_ID)
  )
```

```{r}
# merging RR with the daily emo df
emaact <-
  left_join(
    act,
    mw_emo_filt_merge,
    by = c("ELS_ID", "triggerdate") 
  ) %>%
  drop_na(efficiency)

emaact %>%
  distinct(ELS_ID)

emaact_all <-
  left_join(
    emaact,
    demoact,
    by = "ELS_ID"
  )

emaact_all %>%
  distinct(ELS_ID)

```

```{r}
emaact_allcent <-
  emaact_all %>%
  dplyr::select(
    -c(ends_with("_mw"), ends_with("_MW"))
  ) %>%
  mutate(
    Sex = factor(Sex),
    COVID_AHC_ACT = factor(COVID_AHC_ACT),
    sumsev_type_z = scale(sumsev_type),
    cmep_total_act_z = scale(cmep_total_act),
    asq_total_act_z = scale(asq_total_act)
    ) %>%
  group_by(Sex) %>%
  mutate(
    tanner_average_for_act_z = scale(tanner_average_for_act), # standardizing tanner average within each sex
    tanner_adrenal_act_z = scale(tanner_adrenal_act), # standardizing tanner adrenal within each sex
    tanner_gonadal_act_z = scale(tanner_gonadal_act), # standardizing tnaner gonadal within each sex
    bmi_at_act_z = scale(BMI_ACT), # standardizing bmi within each sex
    tanner_age_for_act_z = scale(tanner_age_for_act) # standardizing age within each sex
  ) %>%
  ungroup() 

```

testing for potential covariates
```{r}
c13 <- lm(scale(efficiency) ~ bmi_at_act_z, data = emaact_allcent)
c13_c <- coeftest(c13, vcov = vcovCL, cluster = ~ELS_ID) # nope

c14 <- lm(scale(efficiency) ~ scale(Household_Income_ACT), data = emaact_allcent)
c14_c <- coeftest(c14, vcov = vcovCL, cluster = ~ELS_ID) # nope

c15 <- lm(scale(efficiency) ~ scale(sumsev_type), data = emaact_allcent)
c15_c <- coeftest(c15, vcov = vcovCL, cluster = ~ELS_ID) # nope

c16<- lm(scale(efficiency) ~ Sex, data = emaact_allcent)
c16_c <- coeftest(c16, vcov = vcovCL, cluster = ~ELS_ID) # nope

c17 <- lm(scale(efficiency) ~ asq_total_act_z, data = emaact_allcent)
c17_c <- coeftest(c17, vcov = vcovCL, cluster = ~ELS_ID) # no
```


```{r}
library(moments)

skewness(emaact_allcent$efficiency) # -1.757371
skewness(emaact_allcent$sleep_time_hrs) # -0.3951045
skewness(emaact_allcent$neg_total) # 0.9783957

emaact_allcent <-
  emaact_allcent %>%
  mutate(
    efficiency_sqrt = sqrt(max(efficiency+1) - efficiency),
    efficiency_log = log10(max(efficiency+1) - efficiency) ,
    sleeptime_sqrt = sqrt(max(efficiency+1) - sleep_time_hrs),
    sleeptime_log = log10(max(efficiency+1) - sleep_time_hrs) ,    
    neg_total_sqrt = sqrt(neg_total),
    neg_total_log = log(neg_total)
  )
skewness(emaact_allcent$efficiency_sqrt) # 0.755437  ehh
skewness(emaact_allcent$efficiency_log) # -0.3352765 better
skewness(emaact_allcent$sleeptime_sqrt) # 0.3677143  better
skewness(emaact_allcent$sleeptime_log) # 0.340609 (choosing the sqrt transform bc direction is closer to raw)
skewness(emaact_allcent$neg_total_sqrt) # -0.02112807  ehh
skewness(emaact_allcent$neg_total_log) # -1.292901 better


m4 <- lm(scale(efficiency_sqrt) ~ scale(neg_total_sqrt) + scale(pos_total), data = emaact_allcent)
summary(m4)
m4_c <- coeftest(m4, vcov = vcovCL, cluster = ~ELS_ID)
confint(m4_c)

m5 <- lm(scale(sleeptime_sqrt) ~ scale(neg_total_sqrt) + scale(pos_total), data = emaact_allcent)
summary(m5)
m5_c <- coeftest(m5, vcov = vcovCL, cluster = ~ELS_ID)
confint(m5_c)
```
